# Part 2: More Experimentation

## On the way to Final Model

After some of our previous trials, we realized that not all factors play an essential role in the college ranking algorithm. To check our understanding, we built a quick linear regression model. From its summary and anova table, we recognized the important factors were the following: *size of undergraduate student*, *SAT score*, *admission rate*, *location* and of course *intercept*.

```{r message=FALSE, warning=FALSE}
LinearModel <- lm(as.numeric(Y2018) ~ UGDS_1617 + LOCALE_collapse + SAT_AVG_1617 + ADM_RATE_1617 + COSTT4_A_1617 + UGDS_WHITE_1617, fullUniversity)
```

```{r message=FALSE, warning=FALSE}
summary(LinearModel)
anova(LinearModel)
```

OMG, what happened to the **cost** variable? Is it because we put the cost variable COSTT4_A_1617 in the second to last position?

In the spirit of scientific experiment (which Kavya and Zuofu happen to have), we switched the order of our linear model and produced another anova test:

```{r message=FALSE, warning=FALSE}
LinearModel2 <- lm(as.numeric(Y2018) ~ UGDS_1617 + LOCALE_collapse + COSTT4_A_1617 + SAT_AVG_1617 + ADM_RATE_1617 + UGDS_WHITE_1617, fullUniversity)
```

```{r message=FALSE, warning=FALSE}
anova(LinearModel2)
```

Certainly, cost remains one of the most important factors for students in the process of college selection. We also acknowledge the significant role that scholarships and financial aid play. However, the order switch experiment sufficiently shows that other variables explain an equal amount of variability.

## What are we thinking?

**"Zuofu was on his way to create the best Bayesian model in the world when his dream got crushed."** --Zuofu

From the above, we made the final decision to incorporate the following variables into our Bayesian model of rankings: size of undergraduate student body, average SAT score, admission rate, and location.

To see how we do with universities and liberal arts colleges, check out our next chapter: **Becoming a proud Bayesian**.


