# Becoming a proud Bayesian

For our finalized Bayesian model, we intend to build predictive models for universities and liberal arts colleges. Then we'll compare the similarities and differences of the two models. (i.e. is some factor important for universities and less so for liberal arts colleges?)

## University Model

### Get some intuition



### Some additional intuition



### Even more intuition

For additional 

### Building the model

```{r}
university_model_3 <- "model{  
    # Data: observations
    for(i in 1:length(y)) {
        y[i] ~ dpois(lambda[i])
        log(lambda[i]) = beta0 + beta1*x1[i] + beta2[z[i]] + beta3*x3[i] + beta4*x4[i]
    }
        # Data: subjects
        beta0 ~ dnorm(8, 1)
        beta1 ~ dnorm(0, 10000)
        beta2[1] <- 0 
        beta2[2] ~ dnorm(0, 0.01)
        beta2[3] ~ dnorm(0, 0.01)
        beta3 ~ dnorm(0, 10)
        beta4 ~ dnorm(0, 0.01)

}"

# COMPILE

y <- as.numeric(fullUniversity$Y2018)

model_data9 <- data.frame(y, x1 = fullUniversity$UGDS_1617, z = as.numeric(fullUniversity$LOCALE_collapse), x3 = fullUniversity$SAT_AVG_1617, x4 = fullUniversity$ADM_RATE_1617)

model_data9 <- na.omit(model_data9)

university_jags_3 <- jags.model(textConnection(university_model_3), 
    data = list(y = model_data9$y, x1 = model_data9$x1, z = factor(model_data9$z), x3 = model_data9$x3, x4 = model_data9$x4),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

# SIMULATE the model
university_sim_3 <- coda.samples(university_jags_3,
    variable.names = c("beta0","beta1","beta2","beta3","beta4"),
    n.iter = 10000)

# STORE the chains in a data frame
university_chains_3 <- data.frame(university_sim_3[[1]])
```


### Model summary

```{r}
summary(university_sim_3)
```


### Posterior inference

For an unknown university located in the city with 10000 undergraduates, student mean SAT score of 1550 and an admission rate of $23\%$ (e.g. Gvictor University), we could predict its ranking from our rjags simulation.

```{r}
university_chains_3 <- university_chains_3 %>%
  mutate(ranking_new = rpois(10000, lambda = exp(beta0 + beta1 * 10000 + beta3 * 1550 + beta4 * 0.23)))

university_chains_3 %>%
  summarize(quantile(ranking_new,0.025),quantile(ranking_new,0.975))
```

The interval is more reflective of the intuitive estimate of Gvictor University's ranking.

## Liberal Arts Colleges

### Get some intuition

Similar to what we do with universities, we get some intuition for our liberal arts college model.

```{r}
summary(lm(full_LiberalArts$Y2018 ~ full_LiberalArts$UGDS_1617 + full_LiberalArts$LOCALE_collapse_lac1 + full_LiberalArts$SAT_AVG_1617 + full_LiberalArts$ADM_RATE_1617))
```

### Building the model


```{r}
liberal_model_3 <- "model{  
    # Data: observations
    for(i in 1:length(y)) {
        y[i] ~ dpois(lambda[i])
        log(lambda[i]) = beta0 + beta1*x1[i] + beta2[z[i]] + beta3*x3[i] + beta4*x4[i]
    }
    # Data: subjects
    beta0 ~ dnorm(0, 0.0001)
    beta1 ~ dnorm(0, 10000)
    beta2[1] <- 0 
    beta2[2] ~ dnorm(0, 0.04)
    beta2[3] ~ dnorm(-10, 0.04)
    beta2[4] ~ dnorm(-15, 0.01)
    beta3 ~ dnorm(-0.3, 100)
    beta4 ~ dnorm(36, 10)

}"

# COMPILE

y <- as.numeric(full_LiberalArts$Y2018)

model_data8 <- data.frame(y, x1 = as.numeric(full_LiberalArts$UGDS_1617), z = as.numeric(full_LiberalArts$LOCALE_collapse_lac1), x3 = as.numeric(full_LiberalArts$SAT_AVG_1617), x4 = as.numeric(full_LiberalArts$ADM_RATE_1617))

model_data8 <- na.omit(model_data8)

liberal_jags_3 <- jags.model(textConnection(liberal_model_3), 
    data = list(y = model_data8$y, x1 = model_data8$x1, z = factor(model_data8$z), x3 = model_data8$x3, x4 = model_data8$x4),
    inits=list(.RNG.name = "base::Wichmann-Hill", .RNG.seed = 454))

# SIMULATE the model
liberal_sim_3 <- coda.samples(liberal_jags_3,
    variable.names = c("beta0","beta1","beta2","beta3","beta4"),
    n.iter = 10000)

# STORE the chains in a data frame
liberal_chains_3 <- data.frame(liberal_sim_3[[1]])
```

### Model summary

```{r}
summary(liberal_sim_3)
```


### Posterior inference

For an unknown liberal arts college in a rural setting with 2000 undergraduates, student mean SAT score of 1400 and an admission rate of $35\%$ (e.g. Hvictor College), we could predict its ranking from our rjags simulation.

```{r}
liberal_chains_3 <- liberal_chains_3 %>%
  mutate(ranking_new = rpois(10000, exp(beta0 + beta1 * 2000 + beta2.4. + beta3 * 1400 + beta4 * 0.35)))

liberal_chains_3 %>%
  summarize(quantile(ranking_new,0.025),quantile(ranking_new,0.975))
```

This is **totally reasonable** and **fantastic**! The interval reflects the intuitive estimate of Hvictor College's ranking.

## Comparison